---
title: "Talking about graphs"
description: |
  Build a basic vocabulary to describe and characterize graphs.
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
set.seed(420)
```

## Talking about graphs

Before we go further into other types of graphs, let's first build a basic vocabulary of how to talk about graphs.
How can we quantify simple things about a graph to characterize certain aspects of it?

### Network size and density

A low-hanging fruit is to first describe the size of a graph.
We can count how many vertices a graph has, this is called the *vertex count* or *node count*.
Unsurprisingly, the number of edges is called the *edge count*.

These two metrics seem inoccuously simple, but they give us everything we need to calculate an interesting metric that gives us a more meaningful way of characterizing a graph: it's *density*.

Look at these two graphs:

```{r}
g1 <- erdos.renyi.game(15, 0.1)
g2 <- erdos.renyi.game(15, 0.9)
plot(g1)
plot(g2)
```

The difference between the two just hits you in the face right away.
But there is a way to quantify it very easily.

So, what is network density?

In essence, it is the actual *edge count* divided by the *number of possible edges* in the graph.

$$
density(G) = \frac{edges(G)}{{possibleedges}(G)}
$$

How many edges can a graph with a fixed node count *n* possibly have?
Well, there would be an edge between any two nodes in the network.
A graph like that is called a *complete graph*.

The number of edges in a complete graph with n nodes is calculated as follows:

$$
\frac{n(n - 1)}{2}
$$

Task: Can you figure out why?



Task: Calculate the network density of the following graph:

```{r}
g <- erdos.renyi.game(10, 0.5)
plot(g)
```



  
Let's now take a look at vertices.
What can we say about a vertex to describe it?

Well, first there is the ID that we gave it, but that's not very descriptive, is it?

In a simple graph like the ones that we just saw, there is one glaring thing that we can describe about a vertex:
The number of edges that attach to it.

```{r}
g <- erdos.renyi.game(7, 0.6)
plot(g)
```

Look at vertex 1.
It has `r degree(g, 1)` edges attached to it.
Vertex 7 has `r degree(g, 7)` edges attached to it.

We can see that this number differs between vertices.
In a small graph like this one, these differences seem insignificant, but you will see that they make a huge difference, especially if you look at larger graphs.

For instance, if you use a graph to model a social network (vertices are people and edges are relationships among them), can you see how much of a difference it would make if one person has 1000 acquaintances and another one has only 20?

The number of edges that attach to a node is called the *degree* of a vertex (or simpler: *node degree*).






















