---
title: "What is a graph?"
description: |
  Develop an intuitive understanding for graphs.
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(networkD3)
set.seed(420)
```

The simplest class of graphs is completely defined by a set of **vertices** and a set of **edges**.
You can think of a vertex as a dot in space and an edge as a line connecting two dots.
In this course, we will refer to vertices as **nodes** because frankly, it's just simpler to say and type.

Let's consider a first example:

```{r}
g <- graph_from_literal(1--3, 1--5, 1--7, 2--4, 2--6, 3--5, 3--6, 4--5, 6--7)
simpleNetwork(as_data_frame(g), charge = -100, nodeColour = "blue")
```

This is a simple graph.
As you can see, each node is labelled with a number which is how we usually keep track of nodes.
Labeling nodes with numbers is not obligatory though.
You can call your nodes whatever you want as long they are **uniquely identifiable** (which is why we call the node labels **IDs**).
Sticking with the number labeling, though, we can define the set of nodes more formally by:

$$
N = \{1, 2, 3, 4, 5, 6, 7\}
$$

Edges are simply described as pairs of nodes: $\{a, b\}, \text{where a, b } \in N$.
Notice that we use the set notation for an edge here because the order of the nodes in an edge does not play a role (Yet! More on that later...).
In this specific case, the set of edges is given by:

$$
E = \{\{1, 3\}, \{1, 5\}, \{1, 7\}, \{2, 4\}, \{2, 6\}, \{3, 5\}, \{3, 6\}, \{4, 5\}, \{6, 7\}\}
$$

In summary, the graph is fully described by:

$$
G = (N, E), \text{where N is the set of nodes and E is the set of edges.}
$$

Now that we have seen our first graph, let's take a look at some other ones.
This is also a graph:

```{r}
g <- graph.ring(7)
simpleNetwork(as_data_frame(g), charge = -100, nodeColour = "blue")
```

Obviously, this one looks a little different.

```{r}
g <- graph.star(7, mode = "undirected")
simpleNetwork(as_data_frame(g), charge = -100, nodeColour = "blue")
```

This one looks yet a little different.

But what differentiates these examples?
[How can we describe graphs in a way that gives us a vocabulary to characterize them in a meaningful way?](./talking-about-graphs.html)

