---
title: "Average Path Length"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(dplyr)
set.seed(420)
```

The **average path length** of a network is the average length of the *shortest path* between any two nodes in the network.
To compute the average path length, we first find the shortest path between all pairs of two nodes and then simply compute the arithmetic mean of these paths.

Let's see this in an example.

```{r}
erdos.renyi.game(5, 0.5) %>% 
  plot(vertex.size = 25)
```


#### Task

When analyzing graphs, it is often helpful to have some basic knowledge of combinatorics.
Can you figure out how many unique pairs of nodes there are for the network above without counting them by hand?


There are ${n\choose k}$ possibilities to choose a unique subset of $k$ elements in a set of $n$ elements.
In the case of our graph above, there are ${5\choose 2}$ ways to select a pair of nodes.

$$
{n\choose k} = \frac{n!}{k!(n-k)!}
$$

Thus, in this case:

$$
{5\choose 2} = \frac{5!}{2!(5-2)!} = \frac{5!}{2!3!} = \frac{5 \cdot 4}{1 \cdot2} = 10
$$

So there are $10$ ways to select a pair of nodes from the node set of the graph.

Let's list them out:

$$
\{1, 2\}, \{1, 3\}, \{1, 4\}, \{1, 5\}, \{2, 3\}, \{2, 4\}, \{2, 5\}, \{3, 4\}, \{3, 5\}, \{4, 5\}
$$

For each of these, we now look for the shortest path between them.

$$
|P_{min}(\{1, 2\})| = 3 \\
|P_{min}(\{1, 3\})| = 1 \\
|P_{min}(\{1, 4\})| = 2 \\
|P_{min}(\{1, 5\})| = 1 \\
|P_{min}(\{2, 3\})| = 3 \\
|P_{min}(\{2, 4\})| = 1 \\
|P_{min}(\{2, 5\})| = 2 \\
|P_{min}(\{3, 4\})| = 2 \\
|P_{min}(\{3, 5\})| = 1 \\
|P_{min}(\{4, 5\})| = 1
$$

Now, we simply take the arithmetic mean of these shortest path lengths:

$$
\overline{P_{min}} = \frac{3 + 1 + 2 + 1 + 3 + 1 + 2 + 2 + 1 + 1}{10} = \frac{16}{10} = 1.6
$$

[ELABORATE]
-> digraphs?
-> weighted graphs?
-> never need to do by hand because of computers























