---
title: "Directed Graphs"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
```


The graphs we've seen so far were **undirected graphs**.
That means simply that the order of nodes in an edge does not matter.
However, what if we restrict this condition?
What if the relationship represented by an edge is **directed**?

That's what **directed graphs** are for!

We draw a directed graph simply by putting an arrow head on one side of an edge:

```{r}
g <- graph_from_literal(1 --+ 2, 1 --+ 3, 2 --+ 4, 2 --+ 5, 3 --+ 7, 6 --+ 2, 6 --+ 7, 7 --+ 1)
plot(g)
```

Why would we need directed graphs?

Consider the social media site Twitter.

How do you connect to people on Twitter?
You follow them.
Following someone on Twitter is a one-sided connection.
If you follow a celebrity, you get their updates on your feed, but they do not get yours in their feed.
If we were to model Twitter as a network graph, we would model this relationship with directed edges.

In contrast, a Facebook friendship is different.
You send someone a friend request and then you are connected both ways.
This relationship could be modeled with undirected edges.

In undirected graphs, edges can be modeled as sets of two nodes because the order of nodes doesn't matter.
Directed graphs, however, cannot be formalized this way because the order of nodes in an edge matters.
This is why we use Tuples to model edges.
The notation for the graph above is the following:

$$
G = (N, E) \\
N = \{1, 2, 3, 4, 5, 6, 7\} \\
E = \{(1, 2), (1, 3), (2, 4), (2, 5), (3, 7), (6, 2), (6, 7), (7, 1)\}
$$






