---
title: "Network Size and Density"
# description: |
#   A new article created using the Distill format.
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(networkD3)
library(dplyr)
set.seed(420)
```


A low-hanging fruit is to first describe the **size** of a graph.
We can count how many nodes and how many edges a graph has (**node count** and **edge count**).

These two metrics seem inoccuously simple, but they give us everything we need to calculate an interesting metric that gives us a more meaningful way of characterizing a graph: its *density*.

Look at these two graphs:

```{r two-example-graphs}
g1 <- erdos.renyi.game(15, 0.3)
g2 <- erdos.renyi.game(15, 0.9)
g1df <- igraph::as_data_frame(g1) %>% 
  rename(source = from, target = to) %>% 
  mutate(source = source - 1, target = target - 1)
g2df <- igraph::as_data_frame(g2) %>% 
  rename(source = from, target = to) %>% 
  mutate(source = source - 1, target = target - 1)
simpleNetwork(g1df, charge = -100, nodeColour = "blue")
simpleNetwork(g2df, charge = -100, nodeColour = "blue")
```

Despite the fact that they both have the same number of nodes, the difference between the two just hits you right between the eyes.
The second one is obviously more *dense* than the first one.

There is an easy way to quantify this.
**Network density** is calculated by dividing the **actual edge count** by the **number of possible edges**, that is, the maximum number of edges possible in a graph with a given node count.

$$
density(G) = \frac{edges(G)}{{edges_{max}}(G)}
$$

The maximum number of edges would be present if there is an edge between each possible pair of nodes.
A graph like that is called a **complete graph**.

The number of edges in a complete graph with $n$ nodes is calculated as follows:

$$
\frac{n(n - 1)}{2}
$$

#### Task

Can you figure out why?

#### Task

Calculate the network density of the following graph:

```{r}
g <- erdos.renyi.game(10, 0.5)
gdf <- igraph::as_data_frame(g) %>% 
  rename(source = from, target = to) %>% 
  mutate(source = source - 1, target = target - 1)
simpleNetwork(gdf, charge = -100, nodeColour = "blue")
```



