---
title: "Node Degree"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(dplyr)
library(networkD3)
set.seed(420)
```


Let's now take a look at vertices.
What can we say about a vertex to describe it?

Well, first there is the ID that we gave it, but that's not very descriptive, is it?

In a simple graph like the ones that we just saw, there is one glaring thing that we can describe about a vertex:
The number of edges that attach to it.

```{r}
g <- erdos.renyi.game(7, 0.6)
g %>% 
  igraph::as_data_frame() %>% 
  rename(source = from, target = to) %>% 
  mutate(source = source - 1, target = target - 1) %>% 
  simpleNetwork(charge = -100, nodeColour = "blue")
```

Look at vertex 2.
It has `r degree(g, 3)` edges attached to it.
Vertex 6 has `r degree(g, 7)` edges attached to it.

We can see that this number differs between vertices.
In a small graph like this one, these differences seem insignificant, but you will see that they make a huge difference, especially if you look at larger graphs.

For instance, if you use a graph to model a social network (vertices are people and edges are relationships among them), can you see how much of a difference it would make if one person has 1000 acquaintances and another one has only 20?

The number of edges that attach to a node is called the *node degree*.